Fast Algorithms for Intersection of Non-matching Grids Using \plucker coordinates.

### 2016-10-01 paper_intersections.tex
Partly reviewed sec 3

%\title{Algorithm for Intersections of Nonmatching Grids of Different Dimensions}
\title{Fast Algorithms for Intersection of Non-matching Grids Using \plucker Coordinates.}

@@

\begin{abstract}
The XFEM and Mortar methods can be used in combination with non-matching or non-conforming grids 
to deal with problems on complex geometries. However the information about the mesh intersection must be provided.
We present algorithms for intersections between 1d and 2d unstructured multi component meshes 
and their intersections with a background unstructured 3d mesh. A common algorithm 
based on the advancing front technique is used for the efficient selection of candidate pairs among 
simplicial elements. Bounding interval hierarchy (BIH) of axes aligned bounding boxes (AABB) of elements
is used to initialize the front tracking algorithm. The family of element intersection algorithms
is built upon a line-triangle intersection algorithm based on the \plucker
coordinates. These algorithms combined with the advancing front technique can reuse the results of calculations 
performed on the neigbouring elements and reduce the number of arithmetic operations. Barycentric coordinates
on each of the intersecting elements are provided for every intersection point. Benchmarks 
of the element intersection algorithms are presented and three variants of the global intersection 
algorithm are compared on the meshes raising from a hydrogeological application.
\end{abstract}

@@

%% main text
\section{Introduction}
The grid intersection algorithms are crucial for several techniques that try to overcome some limitations of the classical finite element method.
The Chimera method \cite{brezzi_analysis_2001}, also called overset grid, and similar Nitche method \cite{massing_efficient_2013}
allow solution of the problems with changing geometry as in the fluid-structure problems. 
The Mortar method \cite{belgacem_mortar_1999} allows domain decomposition, independent meshing of domains, and supports sliding boundaries.
However our primal motivation is usage of XFEM methods \cite{fries_extended/generalized_2010} and non-matching meshes of mixed dimension in groundwater models.

The realistic models of groundwater processes including the transport processes and geomechanics have to deal with 
a complex nature of geological formations including fractures and wells. Although of a small scale, these features may have significant impact 
on the global behavior of the system and their representation in the numerical model is imperative.
One possible approach is to model fractures and wells as lower dimensional objects and introduce their coupling with the surrounding continuum. 
The discretization then leads to the meshes of mixed dimensions, i.e. composed of elements of different dimension. This approach 
called mixed-dimensional analysis in the mechanics \cite{bournival_mesh-geometry_2008} is also studied in the groundwater context, see e.g. 
\cite{martin_modeling_2005}, \cite{fumagalli_numerical_2011}, \cite{brezina_analysis_2015} and 
already adopted by some groundwater simulation software, e.g FeFlow \cite{trefry_feflow:_2007} and Flow123d \cite{flow123d}.
Nevertheless as the complexity of the geometry increase (e.g. when lots of fractures are randomly generated) the compatible meshing becomes painful or even 
impossible. In order to avoid these difficulties we may discretize the continuum and every fracture and well independently, getting a non-matching 
(or incompatible) mesh of mixed dimensions and then apply XFEM to represent jumps of the solution on the fractures or singularities 
around the wells. The prerequisite for such approach is a fast and robust algorithm for calculating intersections of individual meshes.

We consider a composed mesh $\mathcal T$ consisting of simplicial meshes $\mathcal T_i$ of dimensions $d_i \in \{1,2,3\}$, $i=1,\dots,N_\mathcal{T}$
in the 3d ambient space. We assume that every mesh $\mathcal T_i$ is a connected set with no self intersection.
Further we assume only single 3d mesh $\mathcal{T}_1$.
The mesh intersection problem is to find all pairs of elements $L\in \mathcal{T}_i$, $K\in \mathcal{T}_j$, $i\ne j$ that have non-empty intersection 
and to compute that intersection. The mesh intersection problem consists of the two parts: First, generate a set of candidate pairs $(K, L)$. 
Second, compute the intersection of a particular pair.

According to our knowledge, there are lots of works using non-matching grids, yet only few of them discuss algorithms how to compute their intersections. 
Gander and Japhet \cite{gander_algorithm_2013} present the PANG algorithm for 2d-2d and 3d-3d intersections that can be used e.g. for mesh overlapping methods.
They use the advancing front technique to get candidate pairs in linear time. The algorithm is part of the DUNE library \cite{bastian_towards_2005}.
Massing, Larson, and Logg \cite{massing_efficient_2013} present an algorithm for 2d-3d intersections as part of their implementation of the Nitche method 
which is part of the Doplhin project \cite{Dolphin}. They use axes aligned bounding boxes of elements (AABB) and bounding interval hierarchy (BIH) 
to get intersection candidate pairs of elements, the GTS library \cite{gtslib} is used for 2d-3d intersections. 
Finally, there is the work of A. Elsheikh and M. Elsheikh \cite{elsheikh_reliable_2012} presenting an algorithm for 2d-2d mesh union operation which includes 
calculation and imprinting of the intersection curves. They exploit binary space partitioning for search of initial intersection and 
the advancing front method for intersection curve tracking.

In this paper we present a new approach to intersection calculation of simplicial elements of different dimensions based on the \plucker coordinates,
further developing the algorithm of Platis and Theoharis \cite{platis_fast_2003} for ray-tetrahedron intersections. Element intersections based on 
\plucker coordinates are combined with the advancing front method which allows us to reuse \plucker coordinates and their products between neigbouring
elements and reduce the number of arithmetic operations.

The paper is organized as follows. In Section \ref{sec:element intersecitons} the algorithms for
1d-2d, 1d-3d and 2d-3d intersections of simplices are described. In Section \ref{sec:front_advancing} we discuss our implementation of the advancing front 
technique and usage of AABB and BIH for its initialization. Finally, in Section \ref{sec:benchmarks}, we present benchmarks and comparison of individual algorithms.

\section{Element Intersections}
\label{sec:element intersecitons}
In this section, we present algorithms for computing intersection of a pair of simplicial elements of a~different dimension in the 3d ambient space. 
In particular we address intersection algorithms for 1d-2d, 1d-3d, 2d-3d pairs of elements. We have implemented the case 2d-2d as well, however the treatment 
of the special cases is quite technical and not fully completed yet. The fundamental idea is to compute intersection of 1d-2d simplices using the 
\plucker coordinates and reduce all other cases to this one.

We denote $S_i$ a simplicial element with $i+1$ vertices (of dimension $i$). We call vertices, edges, faces and simplices itself the \nfaces and we denote
$M_i$ the set of all \nfaces of the simplex $S_i$.
In general, an intersection can be a point, a line segment or a polygon called \emph{intersection polygon} (IP) in common.
The intersection polygon is represented as a list of its corners called \emph{intersection corners} (IC). The IP data structure keeps also 
reference to the intersecting simplices. A data structure of a single IC consists of:
\begin{itemize}
 \item the barycentric coordinate $\vc w_K$ of IC on $K$,
 \item the dimension $d_K$ of the most specific \nface the IC lies on,
 \item the local index $i_K$ of that \nface on $K$,
\end{itemize}
for each intersecting element $K$ of the pair. The pair $\tau_K = (d_K, i_K)$
is called the topological position of the IC on $K$.

\subsection{\plucker Coordinates}
\plucker coordinates represent a line in 3d space.
Considering a line $p$, given by a point $\vc A$ and its directional vector $\vc{u}$, 
the \plucker coordinates of $p$ are defined as
\[ \pi_p = (\vc{u}_p, \vc{v}_p) = (\vc{u}_p, \vc{u}_p\times \vc A). \]
Further we use a permuted inner product
\[\pi_p \odot \pi_q = \vc{u}_p\cdot \vc{v}_q + \vc{u}_q \cdot \vc{v}_p. \]

@@

\subsection{Intersection Line-Triangle (1d-2d)}
Let us consider a line segment $p$ with parametric equation 
\begin{equation}
    \label{eq:line_parametric}
    \vc X = \vc A + t\vc u,\ t\in (0,1)
\end{equation}
and a triangle $T$ given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. 
\begin{lemma}
The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same non-zero sign if and only if there
is an intersection point $\vc X$ on the $p$ and inside the triangle $T$. 
The barycentric coordinates of $\vc X$ on $T$ are
\begin{equation}
  \label{eq:bary_centric}
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }.
\end{equation}
\end{lemma}
\begin{proof}
Using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ has \plucker coordinates $(\vc u, \vc u \times \vc X)$ since these are invariant to change of the initial point. 
Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
The point $\vc X$ is inside of $T$ if and only if $w_i>0$ for all $i=0,1,2$.
\end{proof}
Having the barycentric coordinates of $\vc X$ on $T$, we can compute also its local coordinate on $p$ from its parametric form:
\begin{equation}
   \label{eq:line}
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\end{equation}
We use $i$ with maximal $|u_i|$ for practical computation.

The calculation of the intersection proceeds as follows:
\begin{enumerate}
    \item Compute or reuse \plucker coordinates and permuted inner products: $\pi_p$, $\pi_i$, $\pi_s \odot \pi_i$, for $i=1,2,3$.
    \item Compute barycentric coordinates $w_i,\ i=1,2,3$ using \eqref{eq:bary_centric}.
    \item If any $w_i$ is less than $\epsilon$, there is no intersection, return empty IP.  
    \item If all $w_i$ are greater than $\epsilon$, we set $\tau_T = (2, 0)$ for the IC.
    \item If one $w_i$ is less than $\epsilon$, intersection is on the edge $s_i$, we set $\tau_T =(1,i)$.
    \item If two $w_i$ are less than $\epsilon$, intersection is at the vertex $V_i$, we set $\tau_T=(0,i)$.
    \item \label{item:coplanar} If all $w_i$ are less than $\epsilon$, the line is coplanar with the triangle, both objects are 
        projected to the plane $x_i=0$ where $i$ is the index of maximal component of the triangle's normal vector.
        Every pair $p$, $s_i$ is checked for an intersection on $T$ boundary either inside $s_i$ or at a vertex $V_i$, setting the topological info $\tau_T$ to 
        $(1, i)$ or $(0, i)$, respectively. At most two ICs are obtained.
    \item For each IC the barycentric coordinates $(1-t, t)$ on the line $p$ are computed according to \eqref{eq:line}. 
    \item If $t\in (-\epsilon, \epsilon)$ or $t\in (1-\epsilon, 1+\epsilon)$, we set $\tau_p = (0,0)$ or $\tau_p = (0,1)$, respectively.
    \item If $t\notin (-\epsilon, 1+\epsilon)$, the IC is eliminated.
\end{enumerate}

Checking the same sign of the inner products can be viewed as a geometric predicate for the presence of the intersection and orientation of the 
line with respect to the triangle. Adaptive-precision evaluation of the geometric predicates was designed by Schewchuk \cite{shewchuk_adaptive_1997}
and used for 2d-2d mesh intersections in \cite{elsheikh_reliable_2012}. However, we rather apply a fixed tolerance check for the zero 
barycentric coordinates and consistently keep the topological positions in this and related algorithms. \noteJB{Can we make the algorithm 
\emph{parsimonious} in the spirit of the Fortune \cite{fortune_stable_1989} quoted by Schewchuk? 
Seems that our problem is more local than the line example that was proven to be NP-hard.}

The algorithms for 1d-3d and 2d-3d intersections use simpler version of the 1d-2d intersection algorithm, in particular the search for ICs in the coplanar case 
(item \ref{item:coplanar}) is not necessary, and the test in the last point is not performed.
%and degenerate cases higher dimensional cases.

%\subsection{Intersection Triangle-Tetrahedron (2d-2d)}
%\noteJB{Just note that this algorithm can also be implemented using 1d-2d, but quite complex to not yet fully tested}

\subsection{Intersection Line-Tetrahedron (1d-3d)}
In this section we consider intersection of a line segment $p$ given by the parametric equation \eqref{eq:line_parametric} with a tetrahedron 
$S_3$. The used algorithm is based on 1d-2d algorithm and closely follows \cite{platis_fast_2003}. Our modification takes into account
intersection with a line segment and consistently propagates topological position of ICs.

@@

  \DontPrintSemicolon
  \SetKw{and}{and}
  \SetKw{continue}{continue}
  \SetKw{break}{break}
  \KwIn{Tetrahedron $S_3$, line segment $p$.}
  \KwOut{List of ICs on sorted along $p$.}  
  $I=\{\}$ 
  \For{unmarked face $f$ of $S_3$}{
    $L$ = intersection($p,\ f$)\;
    \lIf{ $L$ is none or degenerate}{\continue}
    \lIf{ $L$ is inside the edge $e$}{
        set $\tau_{S_4} = (1,e)$\;
        mark faces coincident with $e$
    }
    \ElseIf{ $L$ is at the vertex $v$}{
        set $\tau_{S_4} = (0,v)$\;
        mark faces coincident with $v$
    }
    append $L$ to $I$
    \lIf{$\abs{I}=2$}{\break} 
  }
  \lIf{$\abs{I}=1$ \and   $I$ is outside of $p$}{ erase $I$}
  \ElseIf{$\abs{I}=2$}{    
    trim intersection with respect to the line segment $p$ \label{line:trimming}
  }
\end{algorithm}
Algorithm \ref{algo:13d} first compute line-face intersections for every face of $S_4$ avoiding duplicate computation of ICs on edges and at vertices and
skipping remaining faces once two ICs are found. Tetrahedron has six edges, so 7 \plucker coordinates and 6 inner products are computed at most. Precomputed 
coordinates and products are passed the 1d-2d algorithm which is performed fro the whole line $p$. After collecting line-tetrahedron ICs,
we do the line segment trimming \ref{line:trimming}. If both ICs are out of the line segment $p$ we eliminate both of them. If one of the ICs if out of $p$
we use the closest end point of the line segment instead and interpolate barycentric coordinates of the IC on $S_4$. The topological positions are updated as well.
The result of the algorithm are zero up to two ICs sorted by the parameter $t$ of the line $p$.

\subsection{Intersection Triangle-Tetrahedron (2d-3d)}
The intersection of a triangle $S_2$ and a tetrahedron $S_3$ is an $n$-side intersection polygon (IP), $n\le 7$. The sides of the polygon
lie either on sides of $S_2$ or on faces of $S_3$. Thus each vertex (IC) of the polygon
can arise either from side-face intersection, or from edge-triangle intersection, or be a vertex of $S_2$.
To get all ICs, we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner products, one for every side-edge pair.
Computation of IP consists of three stages: calculation of side-tetrahedron ICs (Section \ref{sec:sides}), 
calculation of edge-triangle ICs (Section~\ref{sec:edges}), oredering of ICs (Section~\ref{sec:ordering}).
The intersection corners are appended to the list $I$ as they are computed however their order on the polygon boundary is 
defined by the \emph{connection tables} $F_g(:)$ and $F_p(:)$. Every side of the polygon that lies on \nface $x\in M_2\cup M_3$ is followed by an IC given by $F_g[x]$ 
and every IC $p$ is followed by the side that lies on $F_p[p]\in M_2\cup M_3$ (see \noteJB{possible figure}). The vertices of the polygon are ordered in the same 
way as is the order and orientation of the sides of $S_2$, that is counterclockwise around the interior with normal pointing to us.
There are special cases when the IP is degenerated to a line or a point. If some IC may possibly be part of the such degenerated IP we put it into the list $J$.

@@

  \uElseIf{ $s$ have 1 degenerate edge $e$}{\nllabel{line:one_degenerated}
    $f$ is face opposite to $e$\;
    \uIf{other two edges $e_a$, $e_b$ have different sign}{
        $z=\edgefaces(e_a)$\;
        replace $g\in z$, $g\ne f$ with $e$, \return $z$
    }\lElse{
        append IC of $v$ to $J$, \return anything
    }
  }  
  \uElseIf{ $s$ have edge $e$ with sign opposite to other two}{ \nllabel{line:different_sign}
    \return edge faces($e$)\;
  }

@@

\section{Advancing Front Method}
\label{sec:front_advancing}
add references...
% BIH:
% http://ainc.de/Research/BIH.pdf
% http://diglib.eg.org/handle/10.2312/EGWR.EGSR06.139-149

Consider now a complex mesh of combined dimensions consisting of \emph{components}, which are sets of connected
elements of the same lower dimension (1d, or 2d), in the space of connected 3d elements, which we shall call a \emph{bulk}. 
Obtaining all of component-bulk intersections is done in two phases: firstly, we look for the first two elements intersecting each other (initialization); secondly, we prolong the intersection by investigating neighbouring elements (intersection tracking).

>>>>

%\title{Algorithm for Intersections of Nonmatching Grids of Different Dimensions}
\title{Fast Algorithms for Intersection of Non-matching Grids Using \plucker coordinates.}

@@

\begin{abstract}
The XFEM and Mortar methods can be used in combination with non-matching or non-conforming grids 
to deal with problems on complex geometries. However the information about the mesh intersection must be provided.
We present algorithms for intersections between 1d and 2d unstructured multi component meshes 
and their intersections a background unstructured 3d mesh. A common algorithm 
based on the advancing front technique is used for the efficient selection of candidate pairs of 
simplicial elements. Bounding interval hierarchy (BIH) of axes aligned bounding boxes (AABB) of elements
is used to initialize the front tracking algorithm. The family of element intersection algorithms
is build upon a line-triangle intersection algorithm based on the \plucker
coordinates. These algorithms combined with the advancing front technique can reuse calculations 
done on the neigbouring elements and reduce the number of arithmetic operations. Barycentric coordinates
on each of the intersecting elements are provided for every intersection point. Benchmarks 
of the element intersection algorithms are provided and three variants of the global intersection 
algorithm are compared on the meshes raising from a hydrogeological application.
\end{abstract}

@@

%% main text
\section{Introduction}
\notePE{How did you decided for 'non-matching' term? Instead of e.g. incompatible or non-conforming?}
\noteJB{Non-matching grid seems to be the most common.}

The grid intersection algorithms are crucial for several techniques that try to overcome some limitations of the classical finite element method.
The Chimera method \cite{brezzi_analysis_2001}, also called overset grid, and similar Nitche method \cite{massing_efficient_2013}
allow solution of the problems with changing geometry as in the fluid-structure problems. 
The Mortar method \cite{belgacem_mortar_1999} allows domain decomposition, independent meshing of domains, and supports sliding boundaries.
However our primal motivation is usage of XFEM methods \cite{fries_extended/generalized_2010} and non-matching meshes of mixed dimension in groundwater models.

The realistic models of groundwater processes including the transport processes and geomechanics have to deal with 
a complex nature of geological formations including the fractures and wells. Although of small scale, these features may have significant impact 
on the global behavior of the system and their representation in the numerical model is imperative.
One possible approach is to model fractures and wells as lower dimensional objects and introduce their coupling with the surrounding continuum. 
The discretization then leads to the meshes of mixed dimensions, i.e. composed of elements of different dimension. This approach 
called mixed-dimensional analysis in the mechanics \cite{bournival_mesh-geometry_2008} is also studied in the groundwater context, see e.g. 
\cite{martin_modeling_2005}, \cite{fumagalli_numerical_2011}, \cite{brezina_analysis_2015} and 
already adopted by some groundwater simulation software, e.g FeFlow \cite{trefry_feflow:_2007} and Flow123d \cite{flow123d}.
Nevertheless as the complexity of the geometry increase (e.g. when lot of fractures are randomly generated) the compatible meshing becomes painful or even 
impossible. In order to avoid these difficulties we may discretize the continuum and every fracture and well independently getting a non-matching 
(or incompatible) mesh of mixed dimensions and then apply XFEM to represent jumps of the solution on the fractures or singularities 
around the wells. The prerequisite for such approach is a fast and robust algorithm for calculating intersections of individual meshes.

We consider a composed mesh $\mathcal T$ consisting of simplicial meshes $\mathcal T_i$ of dimensions $d_i \in \{1,2,3\}$, $i=1,\dots,N_\mathcal{T}$
in the 3d ambient space. We assume that every mesh $\mathcal T_i$ is a connected set with no self intersection.
Further we assume only single 3d mesh $\mathcal{T}_1$.
The mesh intersection problem is to find all pairs of elements $L\in \mathcal{T}_i$, $K\in \mathcal{T}_j$, $i\ne j$ that have non-empty intersection 
and to compute that intersection. The mesh intersection problem consists of the two parts: First, generate a set of candidate pairs $(K, L)$. 
Second, compute the intersection for particular pair.

According to our knowledge there are lot of works using non-matching grids yet only few of them discuss algorithms how to compute their intersections. 
Gander and Japhet \cite{gander_algorithm_2013} present the PANG algorithm for 2d-2d and 3d-3d intersections that can be used e.g. for mesh overlapping methods.
They use the advancing front technique to get candidate pairs in linear time. The algorithm is part of the DUNE library \cite{bastian_towards_2005}.
Massing, Larson, and Logg \cite{massing_efficient_2013} present an algorithm for 2d-3d intersections as part of their implementation of the Nitche method 
which is part of the Doplhin project \cite{Dolphin}. They use axes aligned bounding boxes of elements (AABB) and bounding interval hierarchy (BIH) 
to get intersection candidate pairs of elements, the GTS library \cite{gtslib} is used for 2d-3d intersections. 
Finally, there is the work of Elsheikh and Elsheikh \cite{elsheikh_reliable_2012} presenting an algorithm for 2d-2d mesh union operation which includes 
calculation and imprinting of the intersection curves. They exploit binary space partitioning for search of initial intersection and 
the advancing front method for intersection curve tracking.

In this paper we present a new approach to calculation intersections of simplicial elements of different dimensions based on the \plucker coordinates
further developing the algorithm of Platis and Theoharis \cite{platis_fast_2003} for ray-tetrahedron intersections. Element intersections based on 
\plucker coordinates are combined with the advancing front method which allows us to reuse \plucker coordinates and their products between neigbouring
elements and reduce the number of arithmetic operations. The paper is organized as follows. In Section \ref{sec:element intersecitons} the algorithms for
1d-2d, 1d-3d and 2d-3d intersections of simplices are described. In Section \ref{sec:front advancing} we discuss our implementation of the advancing front 
technique and usage of AABB and BIH for its initialization. Finally, in Section \ref{sec:benchmarks}, we provide benchmarks and comparison of individual algorithms.

\section{Element Intersections}
\label{sec:element intersecitons}
In this section, we present algorithms for computing intersection of a pair of simplicial elements of a~different dimension in the 3D ambient space. 
In particular we address intersection algorithms for 1D-2D, 1D-3D, 2D-3D pairs of elements. We have implemented the case 2D-2D as well however the treatment 
of the special cases is quite technical and not fully completed yet. The fundamental idea is to compute intersection of 1D-2D simplices using the 
\plucker coordinates and reduce all other cases to this one.

We denote $S_i$ a simplicial element with $i+1$ vertices (of dimension $i$). We call vertices, edges, faces and simlices itself the \nfaces and we denote
$M_i$ the set of all \nfaces of the simplex $S_i$.
In general, an intersection can be a point, a line segment or a polygon called \emph{intersection polygon} (IP) in common.
The intersection polygon is represented as a list of its corners called \emph{intersection corners} (IC). The IP data structure keeps also 
reference to the intersecting simplices. A data structure of a single IC consists of:
\begin{itemize}
 \item the barycentric coordinate $\vc w_K$ of IC on $K$,
 \item the dimension $d_K$ of the most specific \nface the IC lies on,
 \item the local index $i_K$ of that \nface on $K$,
\end{itemize}
for each intersecting element $K$ of the pair. The pair $\tau_K = (d_K, i_K)$
we call the topological position of the IC on $K$.

\subsection{\plucker Coordinates}
\plucker coordinates represent a line in 3D space.
Considering a line $p$, given by a point $\vc A$ and its directional vector $\vc{u}$, 
the \plucker coordinates of $p$ are defined as
\[ \pi_p = (\vc{u}_p, \vc{v}_p) = (\vc{u}_p, \vc{u}_p\times A). \]
Further we use a permuted inner product
\[\pi_p \odot \pi_q = \vc{u}_p\cdot \vc{v}_q + \vc{u}_q \cdot \vc{v}_p. \]

@@

\subsection{Intersection Line-Triangle (1D-2D)}
Let us consider a line segment $p$ with parametric equation 
\begin{equation}
    \label{eq:line_parametric}
    \vc X = \vc A + t\vc u,\ t\in (0,1)
\end{equation}
and a triangle $T$ given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$. 
\begin{lemma}
The permuted inner products $\pi_p \odot \pi_{s_i},\, i=0,1,2$ have the same non-zero sign if and only if there
is an intersection point $X$ on the $p$ and inside the triangle $T$. 
The barycentric coordinates of $X$ on $T$ are
\begin{equation}
  \label{eq:bary_centric}
   w_i = \frac{\pi_p \odot \pi_{s_i}}{ \sum_{j=0}^{2} \pi_p \odot \pi_{s_i} }.
\end{equation}
\end{lemma}
\begin{proof}
Using the barycentric coordinates the intersection point can be expressed as $\vc X = \vc V_0 + w_1 \vc s_2 - w_2 \vc s_1$.
The line $p$ have \plucker coordinates $(\vc u, \vc u \times \vc X)$ since these are invariant to change of the initial point. 
Combining these two expressions we get
\[
   \pi_p \odot \pi_{s_1} = \vc u \cdot (\vc s_1 \times \vc V_2) + \vc s_1 \cdot ( \vc u \times [\vc V_0 + w_1\vc s_2 - w_2 \vc s_1])
   =-w_1 \vc u \cdot (\vc s_1 \times \vc s_2).
\]
Since $\vc s_0 + \vc s_1 + \vc s_2=0$ we have $ \vc s_1 \times \vc s_2 = \vc s_2 \times \vc s_0 = \vc s_0 \times \vc s_1$ and thus
\[
   \pi_p \odot \pi_{s_i} = -w_i \vc u \cdot (\vc v_1 \times \vc v_2).
\]
The point $X$ is inside of $T$ if and only if $w_i>0$ for all $i=0,1,2$.
\end{proof}
Having the barycentric coordinates of $X$ on $T$, we can compute also its local coordinate on $p$ from its parametric form:
\begin{equation}
   \label{eq:line}
   X_i  = A_i + t u_i, \text{ for } i=1,2,3
\end{equation}
We use $i$ with maximal $|u_i|$ for practical computation.

The calculation of the intersection proceeds as follows:
\begin{enumerate}
    \item Compute or reuse \plucker coordinates and permuted inner products: $\pi_p$, $\pi_i$, $\pi_s \odot \pi_i$, for $i=1,2,3$.
    \item Compute barycentric coordinates $w_i,\ i=1,2,3$ using \eqref{eq:bary_centric}.
    \item If any $w_i$ is less then $\epsilon$ there is no intersection, return empty IP.  
    \item If all $w_i$ are greater then $\epsilon$, we set $\tau_T = (2, 0)$ for the IC.
    \item If one $w_i$ is less then $\epsilon$, intersection on edge $s_i$, we set $\tau_T =(1,i)$.
    \item If two $w_i$ are less then $\epsilon$, intersection in the vertex $V_i$, we set $\tau_T=(0,,i)$.
    \item \label{item:coplanar} If all $w_i$ are less then $\epsilon$, the line is coplanar with the triangle, both objects are 
        projected to the plane $x_i=0$ where $i$ is the index of maximal component of the triangle's normal vector.
        Every pair $p$, $s_i$ is checked for an intersection on $T$ boundary either inside $s_i$ or in a vertex $V_i$ setting the topological info $\tau_T$ to 
        $(1, i)$ or $(0, i)$ respectively. At most two ICs are obtained.
    \item For each IC the barycentric coordinates $(1-t, t)$ on the line $p$ are computed according to \eqref{eq:line}. 
    \item If $t\in (-\epsilon, \epsilon)$ or $t\in (1-\epsilon, 1+\epsilon)$, we set $\tau_p = (0,0)$ or $\tau_p = (0,1)$, respectively.
    \item If $t\notin (-\epsilon, 1+\epsilon)$, the IC is eliminated.
\end{enumerate}

The check for the same sign of the inner products can be viewed as a geometric predicate for the presence of the intersection and orientation of the 
line with respect to the triangle. Adaptive-precision evaluation of the geometric predicates was designed by Schewchuk \cite{shewchuk_adaptive_1997}
and used for 2d-2d mesh intersections in \cite{elsheikh_reliable_2012}. However, we rather apply a fixed tolerance check for the zero 
barycentric coordinates and consistently keep the topological positions in this and related algorithms. \noteJB{Can we make the algorithm 
\emph{parsimonious} in the spirit of the Fortune \cite{fortune_stable_1989} quoted by Schewchuk? 
Seems that our problem is more local than the line example that was proven to be NP-hard.}

The algorithms for 1d-3d and 2d-3d intersections use simpler version of the 1d-2d intersection algorithm, in particular the search for ICs in the coplanar case 
(item \ref{item:coplanar}) is not necessary and the test in the last point is not performed.
and degenerate cases higher dimensional cases.

%\subsection{Intersection Triangle-Tetrahedron (2D-2D)}
%\noteJB{Just note that this algorithm can also be implemented using 1d-2d, but quite complex to not yet fully tested}

\subsection{Intersection Line-Teraherdon (1D-3D)}
In this section we consider intersection of a line segment $p$ given by the parametric equation \eqref{eq:line_parametric} with a tetrahedron 
$S_3$. The used algorithm is based on 1d-2d algorithm and closely follows \cite{platis_fast_2003}. Our modification takes into account
intersection with a line segment and consistently propagates topological position of ICs.

@@

  \DontPrintSemicolon
  \SetKw{and}{and}
  \SetKw{continue}{continue}
  \SetKw{break}{break}
  \KwIn{Tetrahedron $S_3$, line segment $p$.}
  \KwOut{List of ICs on sorted along $p$.}  
  $I=\{\}$ 
  \For{unmarked face $f$ of $S_3$}{
    $L$ = intersection($p,\ f$)\;
    \lIf{ $L$ is none or degenerate}{\continue}
    \lIf{ $L$ is inside the edge $e$}{
        set $\tau_{S_4} = (1,e)$\;
        mark faces coincident with $e$
    }
    \ElseIf{ $L$ is at the vertex $v$}{
        set $\tau_{S_4} = (0,v)$\;
        mark faces coincident with $v$
    }
    append $L$ to $I$
    \lIf{$\abs{I}=2$}{\break} 
  }
  \lIf{$\abs{I}=1$ \and   $I$ is outside of $p$}{ erase $I$}
  \ElseIf{$\abs{I}=2$}{    
    trim intersection with respect to the line segment $p$ \label{line:trimming}
  }
\end{algorithm}
Algorithm \ref{algo:13d} first compute line-face intersections for every face of $S_4$ avoiding duplicate computation of ICs on edges and at vertices and
skipping remaining faces once two ICs are found. Tetrahedron has six edges, so 7 \plucker coordinates and 6 inner products are computed at most. Precomputed 
coordinates and products are passed the 1D-2D algorithm which is performed fro the whole line $p$. After collecting line-tetrahedron ICs,
we do the line segment trimming \ref{line:trimming}. If both ICs are out of the line segment $p$ we eliminate both of them. If one of the ICs if out of $p$
we use the closest end point of the line segment instead and interpolate barycentric coordinates of the IC on $S_4$. The topological positions are updated as well.
The result of the algorithm are zero up to two ICs sorted by the parameter $t$ of the line $p$.

\subsection{Intersection Triangle-Tetrahedron (2D-3D)}
The intersection of a triangle $S_2$ and a tetrahedron $S_3$ is an $n$-side intersection polygon (IP), $n\le 7$. The sides of the polygon
lie either on sides of $S_2$ or on faces of $S_3$. Thus each vertex (IC) of the polygon
can arise either from side-face intersection, or from edge-triangle intersection, or be a vertex of $S_2$.
To get all ICs, we have to compute at most 12 side-face intersections and at most 6 edge-triangle intersections. However,
to this end we only need to compute 9 \plucker coordinates (3 sides, 6 edges) and 18 permuted inner produts, one for every side-edge pair.
Computation of IP consists of three stages: calculation of side-tetrahedron ICs (Section \ref{sec:sides}), 
calculation of edge-triangle ICs (Section~\ref{sec:edges}), oredering of ICs (Section~\ref{sec:ordering}).
The intersection corners are appended to the list $I$ as they are computed however their order on the polygon boundary is 
defined by the \emph{connection tables} $F_g(:)$ and $F_p(:)$. Every side of the polygon that lies on \nface $x\in M_2\cup M_3$ is followed by an IC given by $F_g[x]$ 
and every IC $p$ is followed by the side that lies on $F_p[p]\in M_2\cup M_3$ (see \noteJB{possible figure}). The vertices of the polygon are oredered in the same 
way as is the order and orientation of the sides of $S_2$, that is counterclockwise around the interior with normal pointing to us.
There are special cases when the IP is degenerated to a line or a point. If some IC may possibly be part of the such degenerated IP we put it into the list $J$.

@@

  \uElseIf{ $s$ have 1 degenerate edge $e$}{\nllabel{line:one_degenerated}
    $f$ is face opposite to $e$\;
    \uIf{other two edges $e_a$, $e_b$ have different sign}{
        $z=\edgefaces(e_a)$\;
        replace $g\in z$, $g\ne f$ with $e$, \return $z$
    }\lElse{
        append IC of $v$ to $J$, \return anything
    }
  }  
  \uElseIf{ $s$ have edge $e$ with sign oposite to other two}{ \nllabel{line:different_sign}
    \return edge faces($e$)\;
  }

@@

\section{Global mesh intersection algorithm}
\noteJB{References already in introduction, possible references to some computer graphics book for AABB and BIH}
Having the algorithms for element-element intersections we can proceed to the mesh intersection algorithm. 
We first compute all component-bulk mesh intersections that is (1d-3d and 2d-3d) and then the 1d-2d and 2d-2d intersections use the 3d mesh
to get the intersection pair candidates.  Element intersections for the particular component and bulk mesh 
are obtained in two phases: firstly, we look for the first two elements intersecting each other (initialization); 
secondly, we prolong the intersection by investigating neighbouring elements (front tracking).
\noteJB{Not sure that "prolong" is the right verb.}

% \textbf{breadth first search} algorithm:
%   \begin{enumerate}
%     \item\label{en:first} Get next unprocessed component element $k$.
%     \item Find intersection candidates $\mathcal K$ in bulk mesh (3D elements).
%     \item \label{en:q2}For $K\in \mathcal K$ compute intersection $(k, K)$.  
%     \item Push the intersection neigbours into queues: $(k, L) \to Q_b$, $(l, K) \to Q_c$. 
%     \item While $(l, L) \in Q_b$ check intersection $(l, L)$. Append queues.
%     \item Pop pair from $Q_c$. (move to the next component element: go to \ref{en:q2}).
%     \item If $Q_c$ is empty, go to \ref{en:first}.
%   \end{enumerate}
